---
title: "Frequentist Concepts"
author: "Keith Allison"
date: "10/21/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Q1: Binomial Probability 1

```{r}
dbinom(3, size = 4 ,prob = 0.75)
```

# Q2: Binomial Probability 2

```{r}
pbinom(3, size = 4 ,prob = 0.75)
```

# Q3: Binomial Probability 3

```{r}
1- pbinom(3, size = 5 ,prob = 0.75)
```

# Q4: Normal Probability 1

```{r}
pnorm(1.2, mean = 2, sd = 2)
```

# Q5: Normal Probability 2

```{r}
1 - pnorm(1.2, mean = 2, sd = 2)
```

# Q6: Normal Probability 3

```{r}
pnorm(3.2, mean = 2, sd = 2) - pnorm(1.2, mean = 2, sd = 2)
```

# Q7: Central Limit Theorem 1

The histogram approaches an approximation of the given beta distribution.


# Q8: Central Limit Theorem 2

The histogram looks like the given beta distribution, but less skewed and much more like a normal distribution.


# Q9: Central Limit Theorem 3

The histogram approaches a normal distribution as I hit the button.

# Q10: Central Limit Theorem 3

When the sample size goes from 1 to 2, the sample size doubles. Since the mean is taken of the samples and follows the central limit theorem, the histogram is normalized instead of being purely a collection of samples from the beta distribution. 

# Q11: Central Limit Theorem 3

The two main factors that determine the width of the sampling distribution of the mean are the standard deviation of the population and the standard error of the sample.

# Q12: Library of Babel 1

```{r}
25*25*25
```
There are 15625 possible 3-character words

# Q13: Library of Babel 2

```{r}
410*40*81 - 410*40*80 
```

$B*25^{16400}$

